---
title: "Class09: Machine Learning mini project"
author: "Barry"
format: html
---

In today's lab we will examine some breast cancer biopsy data and apply our clustering and PCA methods to see what we can learn.

```{r}
wisc.df <- read.csv("WisconsinCancer.csv", row.names = 1)
head(wisc.df)
```

```{r}
skimr::skim(wisc.df)
```

Store diagnosis column for later use we will exclude this from our dataset for analysis

```{r}
# Create diagnosis vector for later and rm this 1st col
diagnosis <- as.factor( wisc.df$diagnosis )
wisc.data <- wisc.df[,-1]
```

> Q1. How many observations are in this dataset?

```{r}
nrow(wisc.df)
```

> Q2. How many of the observations have a malignant diagnosis?

```{r}
sum(wisc.df$diagnosis=="M")
```

```{r}
table(wisc.df$diagnosis)
```

> Q3. How many variables/features in the data are suffixed with _mean?

```{r}
length( grep("_mean", colnames(wisc.df), value=T) )

#colnames(wisc.df)
```

# 2. Principal Component Analysis

We need to use `scale=TRUE` here as shown above with our `skim()` report. We could also look at the sd and mean of our columns and see they are on very different scales.

```{r}
wisc.pr <- prcomp(wisc.data , scale=TRUE )
summary(wisc.pr)

```

Our first PC plot of PC1 vs PC2 colored by the expert diagnosis...

```{r}
#wisc.pr$x
plot(wisc.pr$x[,1], wisc.pr$x[,2], col=as.factor(diagnosis))
```

> Q4. From your results, what proportion of the original variance is captured by the first principal components (PC1)?


```{r}
v <- summary(wisc.pr)
pcvar<- v$importance[3,]
pcvar["PC1"]
```
```{r}
# How many PCs to get 0.7 or more
which(pcvar >= 0.7)[1]

```


```{r}
v <- summary(wisc.pr)
pcvar <- v$importance[3,]
pcvar["PC1"]
```



> Q5. How many principal components (PCs) are required to describe at least 70% of the original variance in the data?

```{r}
which(pcvar > 0.7)[1]
```

> Q6. How many principal components (PCs) are required to describe at least 90% of the original variance in the data?

```{r}
which(pcvar > 0.9)[1]
```

# Clustering in PC space

I will pick 3 PCs here for further analysis but you can use more (e.g. include 90% variance etc.). It is your choice here.

```{r}
d.pc <- dist( wisc.pr$x[,1:3] )
wisc.pr.hc <- hclust(d.pc, method="ward.D2")
```

```{r}
plot(wisc.pr.hc)
```

```{r}
grps <- cutree(wisc.pr.hc, k=2)
table(grps)
```

```{r}
table(diagnosis)
```

```{r}
table(diagnosis, grps)
```

```{r}
plot(wisc.pr$x[,1], wisc.pr$x[,2], col=grps)
```
And a 3D plot

```{r}
library(rgl)

plot3d(wisc.pr$x[,1:3], xlab="PC 1", ylab="PC 2", zlab="PC 3", cex=1.5, size=1, type="s", col=grps)

```



