---
title: "Class 5: Data Visualization with ggplot"
author: "Barry (PID: 911)"
format: pdf
---

## Using GGPLOT

To use ggplot2 we first need to install it on our computers. To do this we will use the function `install.packages()`.

Before I use any package functions I have to load them up with a `library()` call, like so:

```{r}
library(ggplot2)
ggplot(cars)
```
```{r}
#| eval: false
head(cars)
```
There is always the "base R" graphics system, i.e. `plot()`

```{r}

plot(cars)
```
To use ggplot I need to spell out at least 3 things:
- data (the stuff I want to plot as a data.frame)
- aesthetics (aes() values - how the data map to the plot).
- geoms (how I want things drawn)

```{r}
ggplot(cars) +
  aes(x=speed, y=dist) +
  geom_point() +
  geom_smooth(method="lm")
```


```{r}
ggplot(cars) + 
  aes(x=speed, y=dist) +
  geom_point() + 
  geom_smooth(method="lm")
```

## Section 6

Adding more layers to a scatter plot

```{r}
ggplot(cars) + 
  aes(x=speed, y=dist) +
  geom_point() + 
  geom_smooth(method="lm") +
  labs(title = "Stopping Distance of some old cars",
       subtitle = "Some caption",
       x="Speed (MPH)",
       y="Distance (ft)",
       caption = "My fun caption") + 
  theme_bw()
```

## Expression data example

```{r}
url <- "https://bioboot.github.io/bimm143_S20/class-material/up_down_expression.txt"
genes <- read.delim(url)
head(genes)
```


Q. How many genes are there?

```{r}
nrow(genes)
```

There are `r nrow(genes)` in this silly dataset.


Q. How many are "up"

```{r}
sum(genes$State == "up")
```

There are `r sum(genes$State == "up")` "UP" regulated genes in this dataset.
