---
title: "Class 08: Halloween Mini Project"
author: "Barry (PID: Barry)"
format: html
---

Today we will examine 538 Halloween "match-up" data.

First I have to get my data into R

```{r}
candy <- read.csv("candy-data.csv", row.names = 1)
head(candy)
```


> Q1. How many different candy types are in this dataset?

There are `r nrow(candy)`.

```{r}
nrow(candy)
```

> Q2. How many fruity candy types are in the dataset?

```{r}
sum(candy$fruity)
```

```{r}
head( candy[ as.logical(candy$fruity), ] )
```

> Q3. What is your favorite candy in the dataset and what is it’s winpercent value?

```{r}
candy["Caramel Apple Pops",]$winpercent
```
> Q4. What is the winpercent value for “Kit Kat”?

> Q5. What is the winpercent value for “Tootsie Roll Snack Bars”?

```{r}
candy["Tootsie Roll Snack Bars",]$winpercent
```
```{r}
skimr::skim(candy)
```

> Q6. Is there any variable/column that looks to be on a different scale to the majority of the other columns in the dataset?

Sure is, winpercent

> Q7. What do you think a zero and one represent for the candy$chocolate column?

> Q8. Plot a histogram of winpercent values


```{r}
hist(candy$winpercent)
```

```{r}
library(ggplot2)

ggplot(candy) +
  aes(winpercent) +
  geom_histogram()

```



> Q9. Is the distribution of winpercent values symmetrical?


> Q10. Is the center of the distribution above or below 50%?

> Q11. On average is chocolate candy higher or lower ranked than fruit candy?

```{r}
choc.ind <- as.logical(candy$chocolate)
fruit.ind <- as.logical(candy$fruity)

choc.win <- candy[choc.ind,]$winpercent
fruit.win <- candy[fruit.ind,]$winpercent
```

```{r}
mean(choc.win)
mean(fruit.win)

```

> Q12. Is this difference statistically significant?

```{r}
t.test(choc.win, fruit.win)

```

> Q13. What are the five least liked candy types in this set?

```{r}
x <- c(5,2,3,6)
sort( x )
sort(x, decreasing = T)
```


```{r}
x
order(x)
```

```{r}
x[ order(x) ]
```


```{r}
y <- c("D", "A", "E")
order(y)
```
```{r}
y[ order(y) ]
```

```{r}
inds <- order(candy$winpercent)
head(candy[inds,], 5)
```

> Q14. What are the top 5 all time favorite candy types out of this set?

```{r}
inds <- order(candy$winpercent, decreasing = T)
head(candy[inds,], 5)

```

> Q15. Make a first barplot of candy ranking based on winpercent values.


```{r}
ggplot(candy) +
  aes(winpercent, rownames(candy)) +
  geom_col()
```


```{r}
ggplot(candy) +
  aes(winpercent, reorder(rownames(candy),winpercent)) +
  geom_col()
  
```

Let's color these candy up by some scheme that means something to us. I will pick a color for all the fruit candy and one for chocolate etc.

```{r}
my_cols=rep("black", nrow(candy))
my_cols[as.logical(candy$chocolate)] = "chocolate"
my_cols[as.logical(candy$bar)] = "brown"
my_cols[as.logical(candy$fruity)] = "pink"

ggplot(candy) +
  aes(winpercent, reorder(rownames(candy),winpercent)) +
  geom_col(bg=my_cols)

```


#  Correlation structure

```{r}
library(corrplot)

cij <- cor(candy)
corrplot(cij)
```

## PCA

```{r}
pca <- prcomp(candy, scale=T)
summary(pca)
```

```{r}
# Make a new data-frame with our PCA results and candy data
my_data <- cbind(candy, pca$x[,1:3])

ggplot(my_data) +
  aes(PC1, PC2, size=(winpercent/100)) +
  geom_point(col=my_cols) +
  theme_bw()


```

And the loadings - how the original variables contribute to our new PCs

```{r}
loadings <- as.data.frame(cbind(candy=rownames(pca$rotation),
                  pca$rotation[,1:3]))

ggplot(loadings) +
  aes(PC1, candy) +
  geom_col()
```

